Exploring probability distributions for bivariate temporal granularities

### 2019-11-27 main.Rmd
acyclic time granularities

\begin{timedom}\label{def:granule}
Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
\end{granule}

@@

```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
knitr::include_graphics("Figs/linear-time.png")
```

@@

**Dicussion:** According to this definition, $G \trianglelefteq H,$ if each granule H(j) is the union of some granules of G.

For example, $Days \trianglelefteq Months$. However, this  relationship is not fully described until we associate it with periodicity. For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping. If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year. Considering leap years and all their exceptions, the period becomes 400 years.

@@

A granularity H which is periodical with respect to G is specified by:
(i) the R sets of indexes of G, ${S_0,...,S_(R-1)}$ describing the granules of H within one period;
(ii) the value of P; 
(iii) the indexes of first and last granules in H, if their value is not infinite. 
Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.

**Dicussion:** This formula can be explained observing that H (j ) is either one of H (0), . . . , H (R - 1) or, for the periodicity of H, is one of these granules “shifted” ahead or behind on the time line of a finite number of periods. If $H (j ) \epsilon {H (0), . . . , H (R - 1)}$, then the formula ‘j mod R’ defines the index (among those in {0,...,R - 1}) of the granule that must be shifted to obtain H (j).The number of periods each granule of G composing H (j mod R) should be shifted is given by $\lfloor j/R \rfloor$.

@@

>>>>

\begin{timedom}\label{def:granule}
Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
\end{timedom}

@@

```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
knitr::include_graphics("Figs/linear-arr.png")
```

@@

**Dicussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.  $G \preceq H$ if every granule in G is a subset of some granule in H. In \autoref{fig:groups-finer}, both $Days \trianglelefteq Weeks$ and $Days \preceq Weeks$ hold true. However, Weekends groups into Holidays but Weekends are not finer than Holidays. On the other hand, Weekends are finer than weeks but do not group into weeks. Consider another example where $Days \trianglelefteq Months$. This relationship is not fully described until we associate it with periodicity. Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored. The period becomes 400 years with the inclusion of leap years and all their exceptions, .

```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships. Weekends groups into Holidays but Weekends are not finer than Holidays. On the other hand, Weekends are finer than weeks but do not group into weeks. Day groups into and are finer than weeks"}
knitr::include_graphics("Figs/groupsandfiner.png")
```

@@

A granularity H which is periodical with respect to G is specified by:
(i) the R sets of indexes of G, ${S_0,...,S_(R-1)}$ describing the granules of H within one period;
(ii) the value of P; 
(iii) the indexes of first and last granules in H, if their value is not infinite. 
Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.

@@

Re-occurances in time can also happen due to completely irregular mapping between linear granularities. Some public and school holidays repeat every year, but the time period on which they appear might vary from year to year. For example, U.S labour day is the first Monday in September and U.S. Memorial Day is the last day in May.

A classic example can be that of Easter.
